Learning to Dispatch: A Reinforcement Learning Framework for Train Dispatch Networks

### 2025-04-24 paper/main.tex
Finish paper

% What is the problem?
The Train Dispatching Problem (TDP) concerns the real-time management of train movements across a rail network, ensuring safe and efficient use of infrastructure. 
This involves deciding when and where trains should move, stop, or wait, based on factors such as schedules, track availability, and priority rules. 
Dispatching decisions must be made continuously and quickly, especially in high-traffic networks, making the problem both operationally critical and computationally challenging.

@@

% Project scope.
This work focuses on the formulation of the Train Dispatching Problem for RL-based approaches, rather than on the algorithm implementation. 
This thesis emphasizes how the problem can be encoded as an RL task using graph structures to represent rail networks, define meaningful states, actions, and rewards, and manage constraints inherent to railway systems. 
The goal is to provide a robust and extensible framework that future researchers and practitioners can build upon when applying RL methods to train dispatching and related infrastructure scheduling problems.
As demonstrated in recent research on applying reinforcement learning to car and pedestrian-level traffic, combing RL and graph structures can yield powerful results, so long as the algorithms and models are capable of modeling the underlying graphs \cite{gnndrl:Devailly_2022}.

% Paper Outline
To provide a thorough understanding of the approach, the remainder of this paper is organized as follows. 
In Section~\ref{sse:background}, we present the foundational background necessary for our formulation, beginning with a formal description of the Train Dispatch Problem, reinforcement learning, and graph neural networks.
These concepts are then reinforced in a discussion of recent research that inspired this work.
Section~\ref{sse:formulation} details the proposed formulation, introducing Train Operation Graphs (\ref{sss:train_ops}), Resource Conflict Graphs (\ref{sss:resource_conflicts}), and the state and action spaces (\ref{sss:state_space}, \ref{sss:action_space}). 
Section~\ref{sse:results} presents the outcomes of implementing the proposed reinforcement learning framework for the Train Dispatching Problem. It highlights the design of the graph-based DQN agent, the reward function employed, and the limitations observed during experimentation.
Finally, Section~\ref{sse:conclusion} preliminary results are showcased from a graph-based Deep Q-Network (DQN) agent followed by future directions for work.

\section{Problem Background}
\label{sse:background}
\subsection{Train Dispatch Problem}
\label{sss:train}
% This should cover the MIP formulation and the general outline of the problem
% This should also cover the DISPLIB data format and all that.
% Reinforcement Learning (RL) has demonstrated significant potential in solving complex, real-world problems, achieving notable successes in domains such as games \cite{drl:rl_games_industry}, robotics \cite{drl:robotics_review}, and supply chain optimization \cite{drl:rl_supply_chain}.
The TDP is a well-known example of a large-scale combinatorial optimization problem, characterized by its complexity and the need for efficient decision-making under constraints. 
This problem becomes especially critical in dense networks with high traffic volume, where suboptimal dispatching can lead to cascading delays, congestion, or even deadlocks.
Traditional approaches typically formulate TDP as a Mixed-Integer Programming (MIP) problem, where binary decision variables represent route allocations, resource occupations (e.g., track segments), and train movements over time. 
Constraints in this formulation encode physical infrastructure limitations, train precedence relations, safety buffers, and scheduling windows \cite{train:LAMORGESE2013559}. 
While MIP solvers can produce high-quality solutions, they are often limited by scalability, especially under tight time constraints or when dealing with real-time disruptions.
Generally, the TDP presents a significant challenge for existing algorithms due to: \begin{itemize} \item Its high-dimensional decision space and tight operational constraints \item The need for fast and adaptive solutions under uncertainty \item The opportunity to drastically improve throughput, punctuality, and network resilience. \end{itemize}

In practice, the TDP is currently managed either by human dispatchers or by static rule-based systems. 
Human expertise plays a crucial role in adapting to disruptions or prioritizing certain services, but manual dispatching lacks scalability and consistency. 
Rule-based systems, while automated, are quite inflexible and often suboptimal when conditions change. 
These limitations open the door for machine learning-based methods, particularly those capable of learning scalable policies from data and interaction.

@@

Given the current lack of RL-based solutions and the standardized dataset provided by the DISPLIB competition, the TDP stands as a compelling application domain for Reinforcement Learning (RL) and Graph-based learning models.

\subsection{Deep Reinforcement Learning}
\label{sss:reinforcement_learning}
% Cover Markov Decision Processes
% As well as DQN
Deep Reinforcement Learning (DRL) is a subfield of machine learning that combines reinforcement learning (RL) with deep learning techniques to solve complex decision-making problems. 
RL focuses on training agents to make sequential decisions by interacting with an environment, receiving feedback in the form of rewards, and learning to maximize cumulative rewards over time. 
Deep learning, on the other hand, enables the representation of high-dimensional data through neural networks, making it possible to handle complex state and action spaces.

@@

In essence, an MDP is a process where the current state contains all the information necessary to determine the optimal action and its associated reward. 
In reinforcement learning, this process is typically modeled as an interaction between an agent and an environment, governed by underlying dynamics. 
At each timestep, the environment provides the current state $s_t$ to the agent, which selects an action $a_t$. 
This action is then applied to the environment, which updates its state to $s_{t+1}$ and computes the corresponding reward $r_{t+1}$. 
This iterative process is illustrated in Figure~\ref{fig:rl_framework}.

@@

MDPs are widely used in reinforcement learning to model environments where agents learn optimal policies through interaction and feedback.

\subsubsection{Deep Q-Network}
Deep Q-Networks (DQN) are a class of reinforcement learning algorithms that approximate the Q-value function using deep neural networks. The Q-value function, $Q(s, a)$, represents the expected cumulative reward of taking action $a$ in state $s$ and following the optimal policy thereafter. The DQN agent learns this function through experience replay and iterative updates.

@@

\begin{itemize}
    \item \textbf{Q-Network:} A neural network that takes the current state $s_t$ as input and outputs Q-values for all possible actions. 
    \item \textbf{Target Network:} A separate network with weights $\theta^-$, used to stabilize training by providing target Q-values. The target network is periodically updated to match the Q-network.
    \item \textbf{Experience Replay:} A buffer that stores past experiences $(s_t, a_t, r_t, s_{t+1})$. During training, mini-batches of experiences are sampled randomly from the buffer to break correlations and improve learning stability.
    \item \textbf{Bellman Equation:} The Q-network is trained to minimize the temporal difference (TD) error, defined as:
    \[
    \delta = \left[ r_t + \gamma \max_{a'} Q(s_{t+1}, a'; \theta^-) - Q(s_t, a_t; \theta) \right],
    \]
    where $\gamma$ is the discount factor.
    \item \textbf{Exploration vs. Exploitation:} The agent balances exploration (choosing new actions to explore the environment) and exploitation (choosing actions to exploit what it already knows) using an $\epsilon$-greedy policy.
\end{itemize}

The training process involves iteratively updating the Q-network's weights $\theta$ using gradient descent to minimize the TD error. 
The learned Q-values guide the agent in selecting actions that maximize long-term rewards.
For simpler tasks, a basic Q-network may suffice (such as a linear regression or heuristic model). 
However, for complex problems like the TDP, more advanced neural network architectures are likely better suited to handle the challenges effectively.

@@

Graph Neural Networks (GNNs) can be thought of as the graph-based counterpart to Convolutional Neural Networks (CNNs) for images. 
GNNs effectively aggregate information from neighboring nodes and edges, enabling each node to build a representation that incorporates both its own features and the context provided by its local graph structure.
They then apply a pooling operation to extract and summarize key patterns and trends from the data.
This comparison is illustrated in Figure~\ref{fig:cnn_vs_gnn}

@@

\subsubsection{Applications in Reinforcement Learning}
GNNs have shown significant promise in reinforcement learning (RL) for tasks involving graph-structured environments. Munikoti et al.~\cite{gnndrl:munikoti2022challengesopportunitiesdeepreinforcement} provide a comprehensive review of how GNNs enhance RL by enabling agents to reason over relational data, such as traffic networks or social graphs. They highlight the ability of GNNs to generalize across varying graph sizes and structures.

Graph Neural Networks (GNNs) have been successfully integrated with reinforcement learning (RL) to address complex graph-structured problems across various domains. 
For example, Almasan et al.~\cite{gnndrl:ALMASAN2022184} applied GNNs to network routing, demonstrating their ability to optimize routing decisions in dynamic environments.
Tackling the problem of influence maximization in social networks, GraMeR~\cite{gnndrl:munikoti2022gramergraphmetareinforcement} leverage GNNs combined with meta-reinforcement learning to learn transferable policies across diverse graph instances.

Devailly et al.~\cite{gnndrl:Devailly_2022} introduced IG-RL, an inductive graph reinforcement learning approach for large-scale traffic signal control. 
By utilizing GNNs, IG-RL efficiently manages massive traffic networks, enabling scalable and adaptive control strategies. 
Similarly, Agasucci et al.~\cite{gnndrl:agasucci} introduced a DRL-based approach to the train dispatching problem, though their method is limited by the specificity of the dataset employed.

@@

The Train Operation Directed Acyclic Graph (DAG) format provides a structured representation of train operations and their constraints. 
Conceptually, each train is modeled as a DAG, called the operations graph, where each node represents a possible operation the train can choose and each edge represents the transitions between operation.
The primary components of the operations graphs are outlined as follows:

@@

@@

\subsubsection{Resources and Constraints}
Resources represent entities (e.g., track sections) that cannot be shared between trains simultaneously. 
Operations requiring the same resource must satisfy the following:
\begin{itemize}
    \item The \textbf{end event} of one operation must precede the \textbf{start event} of the next.
    \item The \textbf{release time} of the resource must be respected.
\end{itemize}

@@

\subsubsection{Objective Function}
The DISPLIB competition defines the objective as minimizing delay costs, calculated as:
\[
v_i = c \cdot \max(0, t - t_{\text{threshold}}) + d \cdot H(t - t_{\text{threshold}})
\]
where $c$ and $d$ are constants, $t$ is the operation's start time, and $H$ is the Heaviside step function \cite{train:DISPLIB2025}.

Alternative objective functions can also be considered, such as minimizing the total time required for all trains to complete their respective exit operations.

\subsection{Resource Conflict Graphs}
\label{sss:resource_conflicts}
To abstract resource labels from individual operations, this formulation introduces a new type of edge to represent resource conflicts between operations. 
These edges connect operations that share the same resource, transforming the disjointed train operation DAGs into a unified graph structure.

@@

    \end{tikzpicture}
    \caption{Example Train Operation DAG with resource and duration constraints.}
    \label{fig:train_rcg}
\end{figure}

@@

In this formulation, the state space is represented as a Directed Acyclic Graph (DAG) composed of unconnected operation graphs. 
Each node in the graph corresponds to an operation and is characterized by a feature vector that captures relevant attributes of the operation. 
The node feature vector includes the following parameters:

\begin{itemize}
    \item $x_1$: Train ID of the operation.
    \item $x_2$: Operation ID of the operation.
    \item $x_3$: Start lower bound of the operation (0 if not present).
    \item $x_4$: Start upper bound of the operation ($10^6$ if not present).
    \item $x_5$: Minimum duration of the operation (0 if not present).
    \item $x_6$: Can pick (1 if the operation has not been picked and can be picked, 0 otherwise).
    \item $x_7$: Currently picked (1 if a train has selected this operation and no successors exist, 0 otherwise).
    \item $x_8$: Picked (1 if the operation has already been picked, 0 otherwise).
    \item $x_9$: Not picked yet (1 if the operation has not been picked but can be picked in the future, 0 otherwise).
    \item $x_{10}$: Feasible (1 if the operation can be picked, 0 otherwise).
    \item $x_{11}$: Cannot be picked (1 if the operation has not been picked and cannot be picked in the future, 0 otherwise).
    \item $x_{12}$: Threshold of the operation (0 if not present).
    \item $x_{13}$: Haversine component of the operation (0 if not present).
    \item $x_{14}$: Linear delay of the operation (0 if not present).
    \item $x_{15}$: Operation start time (0 if not started yet).
    \item $x_{16}$: Operation end time (0 if not started yet).
\end{itemize}

@@

This representation provides a simple yet comprehensive structure for encoding the operations and their relationships, which can be processed by a Graph Neural Network (GNN) to generate a graph embedding for downstream decision-making tasks.

@@

Each action is represented as a discrete choice, and the agent's policy maps the current state of the environment to one of these actions. The design of this action space ensures that the agent has the flexibility to make scheduling decisions dynamically while adhering to the constraints of the problem.

@@

\subsubsection{Graph Reinforcement Learning Algorithm Development}
The integration of Graph Neural Networks (GNNs) with reinforcement learning (RL) has shown promise in handling graph-structured data. 
However, the development of specialized graph RL algorithms tailored to the unique characteristics of train dispatching remains an open challenge. 
Future research could focus on designing algorithms that leverage the structural properties of train operation and resource conflict graphs, enabling more efficient learning and better generalization across different network configurations.

@@

As a formulation-focused study, this thesis aims to inspire future research in applying reinforcement learning to complex scheduling and resource allocation problems. 
Future efforts should prioritize refining the action space, developing tailored graph reinforcement learning algorithms, and exploring applications in other domains. 
These advancements will build upon the groundwork laid here, contributing to the broader understanding and development of intelligent transportation systems.

>>>>

% What is the problem?
The Train Dispatching Problem (TDP), widely recognized as a classic and challenging problem in Operations Research (OR), concerns the real-time management of train movements across a rail network.
This involves deciding when and where trains should move, stop, or wait, based on factors such as schedules, track availability, and priority rules. 
Dispatching decisions must be made continuously and quickly, especially in high-traffic networks, making the problem both operationally critical and computationally challenging.

@@

% Project scope.
This work focuses on the formulation of the Train Dispatching Problem for RL-based approaches, rather than on the algorithm implementation. 
This thesis emphasizes how the problem can be encoded as an RL task using graph structures to represent rail networks and their many constraints. 
The goal is to provide a robust and extensible framework that future researchers and practitioners can build upon when applying RL methods to train dispatching and related infrastructure scheduling problems.
As demonstrated in recent research on applying RL to car and pedestrian-level traffic, combing RL and graph structures can yield powerful results, so long as the algorithms and models are capable of interpreting the underlying graphs \cite{gnndrl:Devailly_2022}.

% Paper Outline
To provide a thorough understanding of the approach, the remainder of this paper is organized as follows. 
In Section~\ref{sse:background}, we present the foundational background necessary for our formulation, beginning with a formal description of the Train Dispatch Problem, Reinforcement Learning, and Graph Neural Networks.
These concepts are then reinforced in a discussion of recent research that inspired this work.
Section~\ref{sse:formulation} details the proposed formulation, introducing Train Operation Graphs (\ref{sss:train_ops}), Resource Conflict Graphs (\ref{sss:resource_conflicts}), and the state and action spaces (\ref{sss:state_space}, \ref{sss:action_space}). 
Section~\ref{sse:results} presents the results of implementing a preliminary graph-based Deep Q-Network (DQN) agent for the Train Dispatching Problem. 
It highlights the design of the graph-based DQN agent, the reward function employed, and the limitations observed during experimentation.
Finally, Section~\ref{sse:conclusion} concludes this thesis and highlights a few future directions for work.

\section{Problem Background}
\label{sse:background}
\subsection{Train Dispatch Problem}
\label{sss:train}
% This should cover the MIP formulation and the general outline of the problem
% This should also cover the DISPLIB data format and all that.
% Reinforcement Learning (RL) has demonstrated significant potential in solving complex, real-world problems, achieving notable successes in domains such as games \cite{drl:rl_games_industry}, robotics \cite{drl:robotics_review}, and supply chain optimization \cite{drl:rl_supply_chain}.
The TDP is a well-known example of a large-scale combinatorial optimization problem, characterized by its complexity and the need for efficient decision-making under constraints. 
This problem becomes especially critical in dense networks with high traffic volume, where suboptimal dispatching can lead to cascading delays, congestion, or even deadlocks.
Traditional approaches typically formulate TDP as a Mixed-Integer Programming (MIP) problem, where binary decision variables represent route allocations, resource occupations (e.g., track segments), and train movements over time \cite{train:doi:10.1287/opre.2014.1327}. 
Constraints in this formulation encode physical infrastructure limitations, train precedence relations, safety buffers, and scheduling windows \cite{train:LAMORGESE2013559}. 
While MIP solvers can produce high-quality solutions, they are often limited by scalability, especially under tight time constraints or when dealing with real-time disruptions \cite{train:Sch_licke_2025}.
Generally, the TDP presents a significant challenge for existing algorithms due to: \begin{itemize} \item Its high-dimensional decision space and tight operational constraints \item The need for fast and adaptive solutions under uncertainty \item The opportunity to drastically improve throughput, punctuality, and network resilience. \end{itemize}

In practice, the TDP is currently managed by human dispatchers using rule-based decision support systems to assist the scheduling process. 
Human expertise plays a crucial role in adapting to disruptions or prioritizing certain services, but manual dispatching lacks scalability and consistency. 
Rule-based systems, while automated, are quite inflexible and often suboptimal when conditions change. 
These limitations open the door for machine learning-based methods, particularly those capable of learning scalable policies from data and interaction.

@@

Given the current lack of RL-based solutions and the standardized dataset provided by the DISPLIB competition, the TDP stands as a compelling application domain for RL and Graph-based learning models.

\subsection{Deep Reinforcement Learning}
\label{sss:reinforcement_learning}
% Cover Markov Decision Processes
% As well as DQN
Deep Reinforcement Learning (DRL) is a subfield of machine learning that combines RL with deep learning techniques to solve complex decision-making problems. 
RL focuses on training agents to make sequential decisions by interacting with an environment, receiving feedback in the form of rewards, and learning to maximize cumulative rewards over time. 
Deep learning, on the other hand, enables the representation of high-dimensional data through neural networks, making it possible to handle complex state and action spaces.

@@

In essence, an MDP is a process where the current state contains all the information necessary to determine the optimal action and its associated reward. 
In RL, this process is typically modeled as an interaction between an agent and an environment, governed by underlying dynamics. 
At each timestep, the environment provides the current state $s_t$ to the agent, which selects an action $a_t$. 
This action is then applied to the environment, which updates its state to $s_{t+1}$ and computes the corresponding reward $r_{t+1}$. 
This iterative process is illustrated in Figure~\ref{fig:rl_framework}.

@@

MDPs are widely used in RL to model environments where agents learn optimal policies through interaction and feedback.

\subsubsection{Deep Q-Network}
Deep Q-Networks (DQN) are a class of RL algorithms that approximate the Q-value function using deep neural networks. 
The Q-value function, $Q(s, a)$, represents the expected cumulative reward of taking action $a$ in state $s$ and following the optimal policy thereafter. 
The DQN agent learns this function through experience replay and iterative updates.

@@

\begin{itemize}
    \item \textbf{Q-Network:} A neural network that takes the current state $s_t$ as input and outputs Q-values for all possible actions. 
    \item \textbf{Target Network:} A separate network with weights $\theta^-$, used to stabilize training by providing target Q-values. The target network is periodically updated to match the Q-network.
    \item \textbf{Experience Replay:} A buffer that stores past experiences $(s_t, a_t, r_t, s_{t+1})$. During training, mini-batches of experiences are sampled randomly from the buffer to break correlations and improve learning stability.
    \item \textbf{Bellman Equation:} The Q-network is trained to minimize the temporal difference (TD) error, defined as:
    \[
    \delta = \left[ r_t + \gamma \max_{a'} Q(s_{t+1}, a'; \theta^-) - Q(s_t, a_t; \theta) \right],
    \]
    where $\gamma$ is the discount factor.
    \item \textbf{Exploration vs. Exploitation:} The agent balances exploration (choosing new actions to explore the environment) and exploitation (choosing actions to exploit what it already knows).
\end{itemize}

The training process involves iteratively updating the Q-network's weights $\theta$ using gradient descent to minimize the Temporal Difference (TD) error. 
The learned Q-values guide the agent in selecting actions that maximize long-term rewards.
For simpler tasks, a basic Q-network may suffice (such as a linear regression or heuristic model). 
However, for complex problems like the TDP, more advanced neural network architectures are likely better suited to handle the challenges effectively.

@@

GNNs can be thought of as the graph-based counterpart to Convolutional Neural Networks (CNNs) for images. 
GNNs effectively aggregate information from neighboring nodes and edges, enabling each node to build a representation that incorporates both its own features and the context provided by its local graph structure.
They then apply a pooling operation to extract and summarize key patterns and trends from the data.
This comparison is illustrated in Figure~\ref{fig:cnn_vs_gnn}

@@

\subsubsection{Applications in Reinforcement Learning}
GNNs have shown significant promise in RL for tasks involving graph-structured environments. 
Munikoti et al.~\cite{gnndrl:munikoti2022challengesopportunitiesdeepreinforcement} provide a comprehensive review of how GNNs can enhance RL by enabling agents to reason over relational data, such as traffic networks, infection models, or social graphs. 
They highlight the ability of GNNs to generalize across varying graph sizes and structures.

GNNs have been successfully integrated with RL to address complex graph-structured problems across various domains. 
For example, Almasan et al.~\cite{gnndrl:ALMASAN2022184} applied GNNs to network routing, demonstrating their ability to optimize routing decisions in dynamic environments.
Tackling the problem of influence maximization in social networks, GraMeR~\cite{gnndrl:munikoti2022gramergraphmetareinforcement} leverage GNNs combined with meta-reinforcement learning to learn transferable policies across diverse graph instances.

Devailly et al.~\cite{gnndrl:Devailly_2022} introduced IG-RL, an inductive graph reinforcement learning approach for large-scale traffic signal control which enables scalable and adaptive control strategies. 
Similarly to this thesis, Agasucci et al.~\cite{gnndrl:agasucci} introduced a DRL-based approach to the train dispatching problem, though their method is limited by the specificity of the dataset employed.

@@

The Train Operation DAG format provides a structured representation of train operations and their constraints. 
Conceptually, each train is modeled as a DAG, called the operations graph, where each node represents a possible operation the train can choose and each edge represents the transitions between operation.
The primary components of the operations graphs are outlined as follows:

@@

\subsubsection{Objective Function}
The DISPLIB competition defines the objective as minimizing delay costs, calculated as:
\[
v_i = c \cdot \max(0, t - t_{\text{threshold}}) + d \cdot H(t - t_{\text{threshold}})
\]
where $c$ and $d$ are constants, $t$ is the operation's start time, and $H$ is the Heaviside step function \cite{train:DISPLIB2025}.
Alternative objective functions can also be considered, such as minimizing the total time required for all trains to complete their respective exit operations.

@@

@@

\subsubsection{Resources and Constraints}
Resources represent entities (e.g., track sections) that cannot be shared between trains simultaneously. 
Operations requiring the same resource must satisfy the following:
\begin{itemize}
    \item The \textbf{end event} of one operation must precede the \textbf{start event} of the next.
    \item The \textbf{release time} of the resource must be respected.
\end{itemize}
To abstract resource labels from individual operations, this formulation introduces a new type of edge to represent resource conflicts between operations. 
\subsection{Resource Conflict Graphs}
\label{sss:resource_conflicts}
A Resource Conflict Graph abstracts resources as edges, rather than treating resources as attributes of individual operations.
These edges connect operations that share the same resource, transforming the disjointed train operation DAGs into a unified graph structure.

@@

    \end{tikzpicture}
    \caption{Example "super graph" with two trains, Train A and Train B. These individual operation graphs are connected by their resource conflicts, signified by red edges.}
    \label{fig:train_rcg}
\end{figure}

@@

In this formulation, the state space is represented as a "super graph" composed of operation graphs connected by their resource conflict edges. 
Each node in the graph corresponds to an operation and is characterized by a feature vector that captures relevant attributes of the operation. 
The node feature vector includes the following parameters:

\begin{itemize}
    \item $x_1$: Train ID of the operation.
    \item $x_2$: Operation ID of the operation.
    \item $x_3$: Start lower bound of the operation (0 if not present).
    \item $x_4$: Start upper bound of the operation ($10^6$ if not present).
    \item $x_5$: Minimum duration of the operation (0 if not present).
    \item $x_6$: Can pick (1 if the operation has not been picked and can be picked, 0 otherwise).
    \item $x_7$: Currently picked (1 if a train is currently handling this operation, 0 otherwise).
    \item $x_8$: Picked (1 if the operation has already been picked, 0 otherwise).
    \item $x_9$: Not picked yet (1 if the operation has not been picked but can be picked in the future, 0 otherwise).
    \item $x_{10}$: Feasible (1 if the operation can be picked and does not violate any constraints, 0 otherwise).
    \item $x_{11}$: Cannot be picked (1 if the operation has not been picked and cannot be picked in the future, 0 otherwise).
    \item $x_{12}$: Threshold of the operation (0 if not present).
    \item $x_{13}$: Haversine component of the operation (0 if not present).
    \item $x_{14}$: Linear delay of the operation (0 if not present).
    \item $x_{15}$: Operation start time (0 if not started yet).
    \item $x_{16}$: Operation end time (0 if not ended yet).
\end{itemize}

@@

This representation provides a simple yet comprehensive structure for encoding the operations and their relationships, which can be processed by a GNN to generate a graph embedding for downstream decision-making tasks.

@@

Each action is represented as a discrete choice, and the agent's policy maps the current state of the environment to one of these actions. 
The design of this action space ensures that the agent has the flexibility to make scheduling decisions while adhering to the constraints of the problem.

@@

\subsubsection{Graph Reinforcement Learning Algorithm Development}
The integration of GNNs with RL has shown promise in handling graph-structured data. 
However, the development of specialized graph RL algorithms tailored to the unique characteristics of train dispatching remains an open challenge. 
Future research could focus on designing algorithms that leverage the structural properties of train operation and resource conflict graphs, enabling more efficient learning and better generalization across different network configurations.

@@

As a formulation-focused study, this thesis aims to inspire future research in applying reinforcement learning to complex scheduling and resource allocation problems. 
Future efforts should prioritize refining the action space, developing tailored graph reinforcement learning algorithms, exploring other applications, and integrating reinforcement learning with mixed integer programming to achieve scalable optimal solutions. 
These advancements will build upon the groundwork laid here, contributing to the broader understanding and development of intelligent transportation systems.
